---
title: "Final Project"
author: "Emily Byrne"
date: "2023-03-08"
output: html_document
---

# Final Project {.tabset}

## Packages {.tabset}
```{r}
library(tidyverse)
library(ggmap)
ggmap::register_google(key = "private")
```

## Dataset {.tabset}

Cincinnati Traffic Crash Reports

source: https://data.cincinnati-oh.gov/safety/Traffic-Crash-Reports-CPD-/rvmt-pkmq

```{r}
crashes = read_csv("Traffic_Crash_Reports__CPD_.csv")
head(crashes)
```

## Descripton of Data Set {.tabset}

### Dimensions and Missing Values {.tabset}

Dimensions 
```{r}
dim(crashes)
```
Total Missing Values 
```{r}
sum(is.na(crashes))
colSums(is.na(crashes))
```

## Visualizations {.tabset}

#### Question 1 {.tabset}

What is the distribution of level 1 severity car crashes in Cincinnati?

```{r}
# get roadmap of Cincinnati area 
cincinnati = get_map(location = "Cincinnati",
                     zoom = 12,
                     source = "google",
                     maptype = "roadmap")

#create heat map of crashes on top of Cincinnati map 
ggmap(cincinnati) +
  geom_point(data = crashes %>% 
               filter(CRASHSEVERITY == "1 - FATAL"),
             aes(x = LONGITUDE_X, y = LATITUDE_X)) +
  geom_density2d(data = crashes %>% 
               filter(CRASHSEVERITY == "1 - FATAL"),
             aes(x = LONGITUDE_X, y = LATITUDE_X),
             color = "purple") +
  geom_density2d_filled(data = crashes %>% 
               filter(CRASHSEVERITY == "1 - FATAL"),
             aes(x = LONGITUDE_X, y = LATITUDE_X),
             alpha = 0.4) +
  labs(title = "Distribution of Level 1 Severity Crashes in Cincinnati",
       x = "Longitude", 
       y = "Latitude",
       fill = "Number of Crashes")
```
- East Central Parkway
- East 5th & East 6th Street 
- Riverbend Concert Venue 
- Ault Park 

#### Question 2 {.tabset}

Which day of the week has experienced the most crashes when the light condition was considered dark? 

```{r}
# Convert light conditions to factor
crashes$LIGHTCONDITIONSPRIMARY <- factor(crashes$LIGHTCONDITIONSPRIMARY)

# Filter data for "dark" light conditions 
dark_crashes <- subset(crashes, LIGHTCONDITIONSPRIMARY %in% c("3 - DARK - LIGHTED ROADWAY", "4 - DARK – ROADWAY NOT LIGHTED", "5 - DARK – UNKNOWN ROADWAY LIGHTING"))

# Convert day of week to factor
dark_crashes$DAYOFWEEK <- factor(dark_crashes$DAYOFWEEK)

# Create a bar plot with facet_wrap
ggplot(dark_crashes, aes(x = LIGHTCONDITIONSPRIMARY, fill = LIGHTCONDITIONSPRIMARY)) +
  geom_bar(width = 0.6) +
  labs(title = "Number of Crashes by Light Conditions",
       x = "Light Conditions",
       y = "Number of Crashes",
       fill = "Light Conditions") +
  theme_minimal() +
  facet_wrap(~ DAYOFWEEK, nrow = 1, scales = "free_x") +
  scale_x_discrete(labels = c("3 - DARK - LIGHTED ROADWAY" = "3", "4 - DARK – ROADWAY NOT LIGHTED" = "4", "5 - DARK – UNKNOWN ROADWAY LIGHTING" = "5"))
```

